---
title: "EDA Diamonds DataSet"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## This R markdown consists of Exploratory data analysis of Diamonds Dataset

Loading the Dataset
```{r Loading Diamonds DataSet}
library(ggplot2)
data("diamonds")
```
Number of Observations
```{r Observations in Diamonds DataSet}
dim(diamonds)
```
Number of Variables
```{r Variables in Diamonds DataSet}
names(diamonds)

```
Ordered Factors in DataSet
```{r Ordered Factors in Diamonds DataSet}
print("cut")
table(diamonds$cut)
print("color")
table(diamonds$color)
print("clarity")
table(diamonds$clarity)

```

Diamonds Dataset Summary
```{r diamonds summmary}
summary(diamonds)
```
Price Histogram
```{r price histogram, echo=FALSE}
qplot(x= price, data = diamonds, color = I('black'), fill = I('blue'), binwidth = 400)
```

Price Summary
```{r Price summmary}
summary(diamonds$price)
```
Number of diamonds costs less than $500
```{r }
#qplot(x= price, data = subset(diamonds, diamonds$price<500), color = I('black'), fill = I('blue'), binwidth = 10)
sum(diamonds$price<500)

```
Number of diamonds costs less than $250
```{r }
#qplot(x= price, data = subset(diamonds, diamonds$price<250), color = I('black'), fill = I('blue'), binwidth = 10)
sum(diamonds$price<250)
```
Number of diamonds costs greater than equalto $15000
```{r }
#qplot(x= price, data = subset(diamonds, diamonds$price>=15000), color = I('black'), fill = I('blue'), binwidth = 10)
sum(diamonds$price>=15000)
```
Exploring the large peak in price Histogram
```{r Cheaper Diamonds}
qplot(x= price, data = diamonds, color = I('black'), fill = I('blue'), binwidth = 20)+
  scale_x_continuous(limits = c(0,800))
 #ggsave('priceHistogram.png')
```
Price Histogram Based on Cut of the Diamond
```{r price histogram based on cut, echo=FALSE}
qplot(x= price, data = diamonds, color = I('black'), fill = I('blue'), binwidth = 400)+
  facet_wrap(~diamonds$cut, ncol =3,  scales = "free")
```
Cut with highest Priced Diamond
```{r}
by(diamonds$price,diamonds$cut,max)
```
Cut with Lowest Priced Diamond
```{r}
by(diamonds$price,diamonds$cut,min)
```
Cut with Lowest Median Priced Diamond
```{r}
by(diamonds$price,diamonds$cut,median)
```
Price Per Carat Histogram
```{r}
qplot(x= price/carat, data = diamonds, color = I('black'), fill = I('blue'))+
  scale_x_log10()+
  facet_wrap(~diamonds$cut, ncol =3,  scales = "free")
```
Boxplots By clarity
```{r}
qplot(geom="boxplot", y = price,x = clarity, data = diamonds)+
  coord_cartesian(ylim = 1:6000)
```
Boxplots By cut
```{r}
qplot(geom="boxplot", y = price,x = cut, data = diamonds)+
  coord_cartesian(ylim = 1:8000)
```
Boxplots By color
```{r}
qplot(geom="boxplot", y = price,x = color, data = diamonds)+
  coord_cartesian(ylim = 1:8000)
```
Price Range for Middle 50% of the diamonds with the color D
```{r}
summary(subset(diamonds,color=="D")$price)
```
IQR of the diamonds with the best color 
```{r}
IQR(subset(diamonds,color=="D")$price)
```
Price Range for Middle 50% of the diamonds with the color J
```{r}
summary(subset(diamonds,color=="J")$price)
```
IQR of the diamonds with the Worst color 
```{r}
IQR(subset(diamonds,color=="J")$price)
```
Price per carat of diamonds across
the different colors of diamonds using boxplots.
```{r}
qplot(geom="boxplot", y = price/carat,x = color, data = diamonds)+
  coord_cartesian(ylim = 1:6000)
```
The Weight of Diamonds vs carat
```{r}
qplot(geom = "freqpoly", x= carat, data = diamonds , bins = 500)+
  scale_x_continuous(limits = c(0,2.5))
```